---
title: "R Notebook"
output: Probabiity Rules
---

To briefly review:

a random experiment is an action where we can't predict its outcome with certainty, like rolling a dice.

an outcome is a possible result of a random experiment, 
and an event can include more than one outcome. Seeing a 1 is a possible outcome of rolling a dice. Seeing an even number is an example of an event since 2, 4, and 6 are all outcomes that meet the definition of the event.

a sample space is the collection of all possible outcomes of a random experiment. The numbers 1 through 6 are the sample space of a dice roll.

a probability can be understood as a numerical measure of chance. You can calculate it by dividing the number of outcomes you want by the total number of outcomes in the sample space.

We'll start this section with a bit of notation. The probability concepts we mentioned above all revolve around the idea of looking at particular items in a collection. These ideas are nicely described in terms of sets. A set is just a collection of distinct objects. That is, we cannot have duplicate objects in a set. The sample space is a great example of a set since its collection of objects are the various outcomes that make it up. We cannot have duplicate outcomes in a sample space, so it qualifies as a set. Taking a six-sided dice, we would represent its sample space as:


outcome = 1,2,3,4,5,6

A set is denoted by a pair of curly brackets that surround each of its items. The sample space is sometimes denoted with the capital Greek letter 
Ω
 (read as "omega"). Changing the notation from above, the sample space of a dice roll can be expressed as:
 
 Ω = 1,2,3,4,5,6
 
 Instructions

We toss a normal coin two times. After list out all of the possible outcomes, assign the number of items in this sample space to the variable coin_toss_omega

```{r}
coin_toss_omega <- 4L

```

For the exercises below consider the six-sided dice roll. We'll define the events
C & D as follows

C: rolling an even number, {2,4,6}
D: rolling a number greater than 3: {4,5,6}


Assume all outcomes have equal chances of occurring.

Instructions

Calculate:

P(C) — assign your answer to p_c.
P(D) - — assign your answer to p_d.


Calculate the probability of getting a number that is either even or greater than 3, 
P(CorD)     Assign your answer to p_c_or_d.

Calculate the probability of getting a number that is either even and greater than 3, 
P(CandD). Assign your answer to p_c_and_d.

```{r}
p_c <- 3/6
p_d <- 3/6
p_c_or_d <- 4/6
p_c_and_d <- 2/6
```

#Union and Intersection of sets


On the last screen, we considered two events at the same time, such as "
C or D" and "C and D". In terms of sets, the event "C or D" would be called the union of 
C and D. In the last screen with the dice roll, we saw that the union of C and D was made up of the outcomes that satisfied the conditions for either C or D.

Likewise, the event "C and D" would be called the intersection of C and D. The outcomes of an intersection must satisfy the conditions for both  and D. By definition, a union of events will usually contain more events than an intersection of events. Using proper set notation, the union of C and D would be denoted as C∪D. The intersection of Cand D would be written as C∩D


Recall that a set can only contain unique elements. This rule applies to unions and intersections as well, so the set resulting from a union will not contain duplicate elements if they appear in both events. Using our example of the dice roll, the events "roll an even number" and "roll a number equal or greater than 3" both contain the number 4. The union of these events would therefore look like 
{2,3,4,5,6} and not {2,3,4,4,5,6}


With intersections, there are often times where there will be no shared outcomes between the events. When this happens, we say that the resulting intersection is the empty set, denoted as 
∅. You can think of the empty set as the set equivalent of the number zero. The empty set is still a set, but it contains no elements. If we took the intersection of the events "roll an odd number" and "roll an even number", we would get the empty set.



Now that we know a bit more about set notation, we can question our probabilities from the last mission more formally:


P(C or D) becomes P(C∪D).P(C and D) becomes P(C ∩ D)
Moving forward, we will use set notation in our courses. Set notation is important because it is the standard notation in the probability literature. If you consult other probability resources to supplement your learning, you'll stand a better chance of understanding the notation. The 
diagram below summarizes the notation that we've learned on this screen.


Instructions
Consider the following sets:
•	M={100,22,1,2}M={100,22,1,2}
•	N={22, car insurance, 2, house insurance}N={22, car insurance, 2, house insurance}
•	O={HHHH, TTTT, TH}O={HHHH, TTTT, TH}
•	P={Hockey, Cycling, Athletics, Swimming}P={Hockey, Cycling, Athletics, Swimming}
Consider the following set operations and their results:
1.	M∪P=∅ M∪P=∅: If you think the result is correct, assign the boolean TRUE to a variable named operation_1, otherwise assign FALSE.
2.	N∩M={22,2}N∩M={22,2}: If you think the result is correct, assign the boolean TRUE to a variable named operation_2, otherwise assign FALSE.
3.	O∪M={HHHH, TTTT, 100, 22, 2}O∪M={HHHH, TTTT, 100, 22, 2}: If you think the result is correct, assign the boolean TRUE to a variable named operation_3, otherwise assign FALSE.
4.	P∩N=∅P∩N=∅: If you think the result is correct, assign the boolean TRUE to a variable named operation_4, otherwise assign FALSE.


```{r}
operation_1 <- FALSE
operation_2 <- TRUE
operation_3 <- FALSE
operation_4 <- TRUE


```

# Mutually Exclusive Events

When there is no intersection between two events, we call them mutually exclusive events. If two events are mutually exclusive, it means that it is a zero probability for them to happen at the same time. We know that when there is no intersection between events, there are also no outcomes in this set. Since there are no outcomes satisfying the condition of the intersection, the probability is zero. Some examples of mutual exclusivity include:



Getting a 5 (event one) and getting a 3 (event two) when we roll a regular six-sided dice — it's impossible to get both a 5 and 3.
A coin lands on heads (event one) and tails (event two) — it's impossible for a coin to land on both heads and tails.



Visualizing mutual exclusivity is easy too. We can see in the diagram above that there is no overlap between events 
A and B, so they have no events in common. This is the same as saying they are mutually exclusive. Mutual exclusivity plays an important role in deciding how we should calculate the probabilities of unions and intersections.

Before we move on, we'll test ourselves on our ability to detect if events are mutually exclusive.



Instructions
We'll use the experiment of picking a random number between 1 and 20. Consider the following events:
•	A=The number chosen is evenA=The number chosen is even
•	B=The number chosen is divisible by 11B=The number chosen is divisible by 11
•	C=The number chosen is divisible by 5C=The number chosen is divisible by 5
With these events in mind, answer the following:
1.	Are AA and BB mutually exclusive? If you think they are, assign the boolean TRUE to a variable named mutual_exclusive_1, otherwise assign FALSE.
2.	Are BB and CC mutually exclusive? If you think they are, assign the bool
ean TRUE to a variable named mutual_exclusive_2, otherwise assign FALSE.
3.	Are AA and CC mutually exclusive? If you think they are, assign the boolean TRUE to a variable named mutual_exclusive_3, otherwise assign FALSE.



```{r}
mutual_exclusive_1 <- TRUE
mutual_exclusive_2 <- TRUE
mutual_exclusive_3 <- FALSE


```

We know how to calculate the probability of a simple event by dividing the number of outcomes in the event by the total number of outcomes. In this mission, we have started to consider unions and intersections, which are made of multiple events. On this screen, we'll learn how to calculate the probabilities of these compound events.
We'll start by considering a fair six-sided dice. We'll define the following events and their union.

•	AA — getting a number divisible by 3
•	BB — getting a number divisible by 5
•	A∪BA∪B — getting a number divisible by either 3 or 5

We know that the sample space of rolling a six-sided dice is Ω={1,2,3,4,5,6}Ω={1,2,3,4,5,6}. Using the conditions for each event, we know that the outcomes satisfying AA and BB are:
•	A={3,6}A={3,6}
•	B={5}B={5}

Notice that AA and BB are also mutually exclusive. Knowing this fact will be an important role in adjusting union and intersection probabilities when this is not the case. Using the equation for probability, we can calculate the probability of A∪BA∪B using the probabilities of the singular events themselves.
P(A)=number of successful outcomes / total number of possible outcomes=26P(A)=number of successful outcomestotal number of possible outcomes=26
P(B)=number of successful outcomes / total number of possible outcomes=16P(B)=number of successful outcomestotal number of possible outcomes=16
P(A ∪ B)=number of successful outcomestotal number of possible outcomes=36P(A ∪ B)=number of successful outcomestotal number of possible outcomes=36
The calculations above demonstrate a simple relationship between individual event probabilities and their union. We can extract a pattern from above into a formula, which is sometimes called the addition rule:
P(A ∪B)=P(A)+P(B)P(A ∪B)=P(A)+P(B)
One way to visualize the probability of an event is the "area" covered by its circle. Remember that this equation is valid for mutually exclusive events though. If we have another look at these events on a Venn Diagram, we can see how this equation works out:


Instructions

Using the addition rule, find the probability of the following events for a playing card deck. Remember that a new deck of playing cards contains 52 cards. The deck contains 4 suits, or distinct symbols, and each suit has 13 playing cards. These 13 cards include: the ace, the numbers 2 through 10 and the three face cards: Jack, Queen and King.

Drawing a card with either hearts or diamonds as its suit. Assign your answer to p_heart_or_diamonds.
Drawing a face card of any suit — assign your answer to p_face_card.

```{r}
p_heart_or_diamonds <- ((13/52) + (13/52))
p_face_card <- 12/52
```

On the last screen, we said that the addition rule could only be used for mutually exclusive events. What if this wasn't the case? We'll explore this with a slightly altered dice-rolling example from the last screen:
•	CC — getting a number divisible by 3
•	DD — getting a number divisible by 2
Listing out the outcomes that satisfy each of the events, we get:
•	C={3,6}C={3,6}
•	D={2,4,6}D={2,4,6}
And finally, we can try adding up the probabilities of CC and DD to get the probability of their union.
P(C)=2/6P(C)=2/6
P(D)=3/6P(D)=3/6
P(C∪D)?=2/6+3/6=5/6?P(C∪D)?=2/6+3/6=5/6?
If we investigate the union of the outcomes of CC and DD, we see that the number 6 appears in both events. By adding the individual probabilities of CC and DD together without considering this, we have accidentally counted for it twice. The union C∪DC∪D is also an event, so its outcomes should also be a set as well. The probability that we calculated is bigger than it should be as a result.
The problem with using the addition rule from the last screen is that it ignores intersections between the two events. Using the area visualization again, we see that the two circles partially cover each other. By adding the two areas together, we include their shared portion twice and overestimate the probability.


With this in mind, we can alter the addition rule formula to account for the shared area:
P(C ∪ D)=P(C)+P(D)−P(C∩ D)P(C ∪ D)=P(C)+P(D)−P(C∩ D)
We added a new term to the addition rule: subtracting C∩DC∩D. This new term corrects for the overcounting we discussed earlier. Applying this new equation to each of the events, we can correctly calculate the probability of the union.

P(C∪D)=2/6+3/6−1/6=4/6P(C∪D)=2/6+3/6−1/6=4/6
Much better! This formula is the general formula for handling the probabilities of unions and intersections of two events. When AA and BB are mutually exclusive, then the probability of their intersection is 0, simplifying the equation to the one we saw in the previous screen. Rearranging this equation also allows us to figure out the probability of intersections as well.

Instructions
An online betting company offers customers the possibility of betting on a variety of games and events (football, tennis, hockey, horse races, car races, etc.). Based on historical data, the company knows the empirical probabilities of the following events:
•	The probability that a new customer's first bet is on football, event FF, is 0.26.
•	The probability that a new customer's first bet is on tennis, event TT, is 0.11.
•	The probability that a new customer's first bet is on both football and tennis, event T∩FT∩F, is 0.03.
•	Find the probability that a new customer's first bet is either on football or tennis. Assign your answer to p_f_or_t. You can't use theoretical probability formula to solve this, so you'll need to make use of the addition rule.

```{r}
p_f_or_t <- 0.26+0.11-0.03
```








